---
title: "5205 Final Code"
output: html_document
date: "2023-04-27"
---


```{r}
df = read.csv('/Users/AstaAnggono/Desktop/Boston Incidents.csv')
```




Research Question 1:How does the month of year, day of week, and hour of day affect the incident rate in Boston? 
```{r}
#library(ggplot2);library(ggthemes);library(gridExtra)  # For plots 
#library(quantmod);library(xts);library(zoo) # For using xts class objects
#library(forecast) # Set of forecasting functions
#library(fpp); library(fpp2) # Datasets from Forecasting text by Rob Hyndman
#library(tseries) # for a statistical test
#library(dplyr) # Data wrangling
```

```{r}
df$DATE <- as.Date(df$DATE)
df <- df[complete.cases(df$DATE ), ]
rownames(df) <- seq(nrow(df))
head(df)
df2 <- cbind(original_row_names = row.names(df), df)
head(df2)
```

```{r}
#crime numbers for month of years
df$DATE2 <- format(df$DATE, "%Y-%m")
head(df)
monthly_crime <- df %>%
  group_by(DATE2) %>%
  summarise(total_incidents = n())

monthly_crime  
```

```{r}
#library(xts)
#library(lubridate)
monthly_crime_xts <- xts(monthly_crime$total_incidents, order.by = as.yearmon(monthly_crime$DATE2))
monthly_crime_xts
monthly_ts <- as.ts(monthly_crime_xts, start = c(year(start(monthly_crime_xts)), month(start(monthly_crime_xts))))
start(monthly_ts)
end(monthly_ts)
```

```{r}
#plot by month of years
monthly_ts %>%
  autoplot()
```

```{r}
#seasonal plot by month of years
ggseasonplot(monthly_ts)
```

```{r}
agg_bostoncrime <- df2 %>%
  group_by(DATE) %>%
  summarize(total_occurrences = n())
head(agg_bostoncrime)
```


```{r}
#crime numbers for hour of day 
summary(df2)
df2$TimeOfDay <- ifelse(df2$HOUR >= 8 & df2$HOUR <= 20, "Day", "Night")
```

```{r}
agg_bostoncrime_daynight <- df2 %>%
  group_by(TimeOfDay) %>%
  summarize(total_occurrences = n())
head(agg_bostoncrime_daynight)
```

```{r}
agg_bostoncrime$dayofweek <-weekdays(agg_bostoncrime$DATE)
agg_bostoncrime
```

```{r}
#crime numbers for day of weeks
agg_bostoncrime2<-agg_bostoncrime%>%
  group_by(dayofweek)%>%
  summarize(sum(total_occurrences))
agg_bostoncrime2
names(agg_bostoncrime2) <- c("DayOfWeek", "TotalOccurrences")
agg_bostoncrime2
```

```{r}
#library(ggplot2)
ggplot(agg_bostoncrime2, aes(x = DayOfWeek, y = TotalOccurrences)) + 
  geom_bar(stat = "identity")+
  xlab('Day of the week') +
  ylab("Total") +
  ggtitle("Total occurrences by day of the week")
```













Research Question 2: Is there a clear trend or seasonal change in the number of incidents and the number of shootings over the past few years?  Is there any significant changes on specific dates (e.g., holidays, hot events, etc.) ?

```{r}
#re-read the dataset
df = read.csv('/Users/wushuyan/Desktop/Boston Incidents.csv')
nrow(df)
```


```{r}
unique(df$SHOOTING)
```

```{r}
# Create a column values for column 'Shooting'
df$SHOOTING_OR_NO <- ifelse(df$SHOOTING == "Shooting reported", 1, 0)
```

```{r}
sum(df$SHOOTING_OR_NO)
```


```{r}
#library(ggplot2);library(ggthemes);library(gridExtra)  # For plots 
#library(quantmod);library(xts);library(zoo) # For using xts class objects
#library(forecast) # Set of forecasting functions
#library(fpp); library(fpp2) # Datasets from Forecasting text by Rob Hyndman
#library(tseries) # for a statistical test
#library(dplyr) # Data wrangling
```


```{r}
df$DATE <- as.Date(df$DATE)
```

```{r}
grouped_df <- df %>%
  group_by(DATE) %>%
  summarize(Crime_Count = n(), Shooting_Count = sum(SHOOTING_OR_NO))
```

```{r}
head(grouped_df)
```

```{r}
grouped_df$DATE <- as.Date(trimws(grouped_df$DATE), format = "%Y-%m-%d")
nrow(grouped_df)
sum(is.na(grouped_df$DATE))
which(is.na(grouped_df$DATE))
grouped_df[which(is.na(grouped_df$DATE)), ]
complete_rows <- complete.cases(grouped_df$DATE)
grouped_df <- grouped_df[complete_rows, ]
nrow(grouped_df)
ts <- data.frame(DATE = seq(min(grouped_df$DATE), max(grouped_df$DATE), by = "day"))
```

```{r}
#see monthly average incident count over time
#library(lubridate)
grouped_df %>%
  mutate(DATE = as.POSIXct(DATE, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(month = floor_date(DATE, "month")) %>%
  group_by(month) %>%
  summarize(avg_Crime_Count = mean(Crime_Count)) %>%
  ggplot(aes(x = month)) +
  geom_line(aes(y = avg_Crime_Count, color = "avg_Crime_Count")) +
  scale_color_manual(values = c("avg_Crime_Count" = "blue")) +
  labs(x = "Month",
       y = "Average Crime Cases Count",
       color = "Event Type",
       title = "Monthly Averages of Incidents")
```

```{r}
#see monthly average incident count with shooting reported over time
#library(lubridate)
grouped_df %>%
  mutate(DATE = as.POSIXct(DATE, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(month = floor_date(DATE, "month")) %>%
  group_by(month) %>%
  summarize(avg_Shooting_Count = mean(Shooting_Count)) %>%
  ggplot(aes(x = month)) +
  geom_line(aes(y = avg_Shooting_Count, color = "avg_Shooting_Count")) +
  scale_color_manual(values = c("avg_Shooting_Count" = "red")) +
  labs(x = "Month",
       y = "Average Shooting Cases Count",
       color = "Event Type",
       title = "Monthly Averages of Shootings")
```

```{r}
#library(tidyverse)

# create a new column for the year and month
grouped_df <- grouped_df %>% 
  mutate(year_month = floor_date(DATE, unit = "month"))
grouped_df

# get the starting and ending dates
start_date <- as.Date(paste(year(min(grouped_df$DATE)), month(min(grouped_df$DATE)), "01", sep = "-"))
end_date <- as.Date(paste(year(max(grouped_df$DATE)), month(max(grouped_df$DATE)), "01", sep = "-")) + months(1) - days(1)

# calculate monthly sums for Shooting cases
monthly_count_Shooting <- grouped_df %>% 
  mutate(year_month = format(DATE, "%Y-%m")) %>% 
  group_by(year_month) %>% 
  summarize(Shooting_Count = sum(Shooting_Count)) %>% 
  ungroup() %>% 
  pull(Shooting_Count)
  
# create the time series data for Shooting Count
ts_df_Shooting <- ts(monthly_count_Shooting, start = c(year(start_date), month(start_date)), end = c(year(end_date), month(end_date)), frequency = 12)
start(ts_df_Shooting)
end(ts_df_Shooting)

# calculate monthly sums for Crime cases
monthly_count_Crime <- grouped_df %>% 
  mutate(year_month = format(DATE, "%Y-%m")) %>% 
  group_by(year_month) %>% 
  summarize(Crime_Count = sum(Crime_Count)) %>% 
  ungroup() %>% 
  pull(Crime_Count)

# create the time series data for Crime Count
ts_df_Crime <- ts(monthly_count_Crime, start = c(year(start_date), month(start_date)), end = c(year(end_date), month(end_date)), frequency = 12)
start(ts_df_Crime)
end(ts_df_Crime)

```

```{r}
ts_df_Shooting
```

```{r}
plot(ts_df_Shooting)
```

```{r}
ts_df_Crime
```

```{r}
plot(ts_df_Crime)
```

```{r}
ts_df_Shooting%>%
  stl(s.window = 'periodic')%>%
  autoplot()
```

```{r}
ts_df_Crime%>%
  stl(s.window = 'periodic')%>%
  autoplot()
```



To see if there's any significant change of Incident occurrences and Shooting incidents occurrences on Holidays
We define Holidays as
#Independence Day
#Labor Day
#martin luther king day
#Halloween
#Thansgiving
#Christmas
#New Year's Eve
```{r}
holidays <- c("2015-07-04", "2015-09-07", "2015-11-26", "2015-10-31", "2015-12-24", "2015-12-25", "2015-12-26", "2015-12-31", 
              "2016-01-18", "2016-07-04", "2016-09-05", "2016-11-24", "2016-10-31", "2016-12-24", "2016-12-25", "2016-12-26", "2016-12-31", 
              "2017-01-16", "2017-07-04", "2017-09-04", "2017-11-23", "2017-10-31", "2017-12-24", "2017-12-25", "2017-12-26", "2017-12-31", 
              "2018-01-15", "2018-07-04", "2018-09-03", "2018-11-22", "2018-10-31", "2018-12-24", "2018-12-25", "2018-12-26", "2018-12-31", 
              "2019-01-21", "2019-07-04", "2019-09-02", "2019-11-28", "2019-10-31", "2019-12-24", "2019-12-25", "2019-12-26", "2019-12-31", 
              "2020-01-20", "2020-07-04", "2020-09-07", "2020-11-26", "2020-10-31", "2020-12-24", "2020-12-25", "2020-12-26", "2020-12-31", 
              "2021-01-18", "2021-07-04", "2021-09-06", "2021-11-25", "2021-10-31", "2021-12-24", "2021-12-25", "2021-12-26", "2021-12-31",
              "2022-01-17", "2022-07-04", "2022-09-05", "2022-11-24", "2022-10-31", "2022-12-24", "2022-12-25", "2022-12-26", "2022-12-31")

```

```{r}
df_holiday <- grouped_df %>% mutate(holiday = if_else(DATE %in% ymd(holidays), 1, 0))
df_holiday<-df_holiday%>%
  filter(holiday==1)
df_holiday
```

```{r}
#Shooting Count during Holidays
df_holiday_Shooting <- df_holiday[, c("DATE", "Shooting_Count")]
df_holiday_Shooting <- left_join(ts,df_holiday_Shooting)
df_holiday_Shooting <- df_holiday_Shooting%>%
  rename(Holiday_Shooting_Count = Shooting_Count)
df_holiday_Shooting
```

```{r}
#Crime Count during Holidays
df_holiday_Crime <- df_holiday[, c("DATE", "Crime_Count")]
df_holiday_Crime <- left_join(ts,df_holiday_Crime)
df_holiday_Crime <- df_holiday_Crime%>%
  rename(Holiday_Crime_Count = Crime_Count)
df_holiday_Crime
```

```{r}
# create the plot
ggplot(df_holiday_Shooting, aes(x = DATE, y = Holiday_Shooting_Count, fill = factor(year(DATE)))) + 
  geom_bar(stat = "identity", width = 0.5, color = "red") +
  theme_minimal() +
  labs(title = "Shooting Count on Holidays and Major Events in Boston",
       x = "Holidays", y = "Count of Shooting Cases")

```

```{r}
# create the plot
ggplot(df_holiday_Crime, aes(x = DATE, y = Holiday_Crime_Count, fill = factor(year(DATE)))) + 
  geom_bar(stat = "identity", width = 0.5, color = "blue") +
  theme_minimal() +
  labs(title = "Crime Count on Holidays and Major Events in Boston",
       x = "Holidays", y = "Count of Crime Cases")
```


```{r}
#library(gridExtra)
#library(ggplot2)
#library(lubridate)

df_holiday_Shooting_view <- ggplot(df_holiday_Shooting, aes(x = DATE, y = Shooting_Count, fill = factor(year(DATE)))) + 
  geom_bar(stat = "identity", width = 0.5, color = "red") +
  theme_minimal() +
  labs(title = "Shooting Count on Holidays and Major Events in Boston",
       x = "Holidays", y = "Count of Shooting Cases")

grouped_df <- grouped_df %>%
  mutate(month = format(as.Date(DATE), "%Y-%m")) 

grouped_df <- grouped_df %>%
  group_by(month) %>%
  mutate(avg_Shooting_Count = mean(Shooting_Count, na.rm = TRUE)) %>%
  ungroup()

grouped_df <- grouped_df %>%
  group_by(month) %>%
  mutate(avg_Crime_Count = mean(Crime_Count, na.rm = TRUE)) %>%
  ungroup()

df_final <- grouped_df %>%
  left_join(df_holiday_Shooting, by = "DATE")%>%
  left_join(df_holiday_Crime, by = "DATE")

df_final_shooting <- df_final[,c('DATE','avg_Shooting_Count','Holiday_Shooting_Count')]
df_final_shooting

df_final_crime <- df_final[,c('DATE','avg_Crime_Count','Holiday_Crime_Count')]
df_final_crime

```

```{r}
ggplot(df_final_shooting) + 
  geom_col(aes(x = DATE, y = Holiday_Shooting_Count), size = 1.5, color = "lightblue", fill = "white") +
  geom_line(aes(x = DATE, y = avg_Shooting_Count), size = 1, color="darkred", group = 1)
```

```{r}
ggplot(df_final_crime) + 
  geom_col(aes(x = DATE, y = Holiday_Crime_Count), size = 1.5, color = "lightgreen", fill = "white") +
  geom_line(aes(x = DATE, y = avg_Crime_Count), size = 1, color="orange", group = 1)
```











Research Question 3: Is there a significant association between the occurrence of shooting incidents and different district locations and hour ranges of incidents?
```{r}
#rerun the dataset
df = read.csv('/Users/wushuyan/Desktop/Boston Incidents.csv')
```

```{r}
# Create a column values for column 'Shooting'
df$SHOOTING_OR_NO <- ifelse(df$SHOOTING == "Shooting reported", 1, 0)
sum(df$SHOOTING_OR_NO)
```

```{r}
#library(caret)

offense_shooting <- df[, c('OFFENSE_CODE_GROUP', 'DISTRICT_NAME', 'HOUR_RANGE', 'SHOOTING_OR_NO')]
head(offense_shooting)

mylogit <- glm(SHOOTING_OR_NO ~ DISTRICT_NAME + HOUR_RANGE, data = offense_shooting, family = "binomial")

summary(mylogit)
# Print the coefficients
```









Research Question 4: Are there certain types of incident that tend to occur together with specific time or location of Boston?
## Read data
```{r}
#library(dplyr)
#library(readxl)
#library(ggplot2)
#library(dplyr)
#library(dplyr)
#library(stringr)
#library(arules)
#library(arulesViz)
#library(RColorBrewer)
#re-read dataset
df=read.csv('/Users/wushuyan/Desktop/Boston Incidents.csv')
df <- df %>% mutate(IND = as.character(DATE))
df$IND <- substring(df$IND, 1, 15)
df$OFFENSE_NAME <- sub(",.*", "", df$OFFENSE_NAME)
df <- group_by(df, IND)
```


## Prepare the Data
```{r}
df = df[complete.cases(df), ]
df$DATE = as.Date(df$DATE,format = "%m/%d/%Y")
df$OFFENSE_NAME = as.factor(df$OFFENSE_NAME)
baskets = df %>%
  arrange(desc(IND))%>%
  mutate(OFFENSE_NAME = str_to_title(OFFENSE_NAME))%>%
  group_by(IND,DATE)%>%
  summarise(OFFENSE_TYPE = paste(OFFENSE_NAME,collapse=','), .groups = 'drop')
baskets$DATE=NULL
baskets$date=NULL
write.csv(baskets,'baskets6.csv',quote = F,row.names = F)
```

## Read Clean Data


```{r}
items = read.transactions('baskets6.csv', format = 'basket', sep=',',skip=1)
items
rules1 = apriori(items,parameter = list(support = 0.001, confidence = 0.05))
```


## Examine top 15 items with support>0.04-only itemsets that appear in at least 4% of the transactions
```{r}
itemFrequencyPlot(items, support = 0.04, cex.names=0.8, 
                  type = "relative", horiz = TRUE, col = "steelblue", las = 1,topN=15,
                  xlab = paste("Proportion of Market Baskets Containing Item",
                               "\n(Item Relative Frequency or Support)"))
```

## Visualize association rules
```{r}
plot(rules1,method='scatterplot',measure=c('support','confidence'),
     control=list(jitter=0, col = rev(brewer.pal(9, "Greens")[4:9])),shading = "lift")
```

## Grouped matrix of rules
```{r, fig.width=10, fig.height=6}
plot(rules1, method="grouped",control=list(col = rev(brewer.pal(9, "Greens")[4:9])))
```

## Visualize plot for relationship between district and offense type
# Delet both "Unspecied" and "Others" reason of crime in the data
```{r}
crime = names(sort(table(df$OFFENSE_CODE_GROUP), decreasing = TRUE))
boston_clean = crime[!(grepl("Unspecified", crime) | grepl("Other", crime))]
df_crime = subset(df, OFFENSE_CODE_GROUP %in% boston_clean)
df_crime
```

## Frequency of crime type with district
```{r}
freq_district_table = table(df_crime$DISTRICT_NAME, df_crime$OFFENSE_CODE_GROUP)
frequency_dis = as.data.frame(freq_district_table)
# Drop the crime type wit low frequency
frequency_dis = frequency_dis[frequency_dis$Freq >= 300, ]
frequency_dis
```
# Finds out what kind of crime are most frequent
```{r}
top_1 <- frequency_dis %>%
  group_by(Var1) %>%
  top_n(1, Freq) %>%
  arrange(Var1, desc(Freq))
top_1
#Larceny exists in almost every district with highest frequency
```

## Heatmap of crime type
```{r heatmap fig.height=8, fig.width=6}
ggplot(frequency_dis, aes(x = Var2, y = Var1)) +
  geom_tile(aes(fill = Freq)) +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "Crime Type", y = "District", fill = "Frequency") +
  ggtitle("Heatmap of Crime Type Frequency by District") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 8))

```


## Find top 10 crime type 
```{r}
top_crime = names(sort(table(df_crime$OFFENSE_CODE_GROUP), decreasing = TRUE))[1:10]
top_crime
```

## Frequency of diiferent crime type in each district
```{r}
freq_table = table(df_crime$DISTRICT_NAME, df_crime$OFFENSE_CODE_GROUP)
frequency = as.data.frame(freq_table)
frequency
```


## Visualize relationship between district and crime type
```{r}
top_crime = frequency %>% group_by(Var1) %>%
top_n(5, Freq) %>%
ungroup() 
```

```{r}
ggplot(top_crime, aes(x=Var1, y=Freq, fill=Var2)) +
geom_bar(stat="identity", position="dodge") +
scale_fill_brewer() +
labs(x="District", y="Frequency", title=" Top 5 Most Frequent Crime in each District ") +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
scale_y_continuous()
```












